\section{Introduction}

Social media platforms have fundamentally reshaped global communication, serving as the primary infrastructure for news dissemination, political discourse, and social interaction. However, this unprecedented connectivity has been aggressively weaponized by \textit{social bots}---automated or semi-automated accounts controlled by algorithmic scripts. The scale of this threat is both measurable and growing: a 2025 Carnegie Mellon University study analyzing approximately 200 million users across seven global events reports that social media activity consists of roughly 20\% bots and 80\% humans, with bot accounts systematically differing from humans in linguistic patterns, interaction structure, and content dissemination behavior.\footnote{\url{https://www.nature.com/articles/s41598-025-96372-1}} The economic consequences are equally alarming. The World Economic Forum ranks disinformation among the top global risks for 2025, while a 2019 study by Professor Roberto Cavazos at the University of Baltimore estimated the annual global financial toll of fake news at approximately \$78 billion, including \$39 billion in stock market losses and \$17 billion in disinformation-driven poor financial decisions; in one illustrative incident, a single fake tweet from a hacked Associated Press account in 2013 briefly erased \$136 billion from the S\&P 500 within three minutes.\footnote{\url{https://www.weforum.org/stories/2025/07/financial-impact-of-disinformation-on-corporations/}} At the consumer level, the U.S. Federal Trade Commission reported that scams originating on social media caused \$2.1 billion in reported losses in 2025---an eightfold increase since 2020---with investment fraud accounting for over \$1.1 billion of that total.\footnote{\url{https://www.ftc.gov/news-events/news/press-releases/2026/04/new-ftc-data-show-people-have-lost-billions\\-social-media-scams}} Operating cohesively as coordinated robotic networks, these adversarial agents are thus deployed at scale to manipulate public opinion, artificially amplify disinformation, and distort financial markets through fake-follower economies. As these bots become increasingly sophisticated, treating them as hostile agents within a complex human-machine interaction graph has elevated bot detection to a critical priority within the broader Artificial Intelligence and Cybersecurity research agendas.

Early detection systems relied primarily on node-level feature engineering, analyzing textual patterns or account metadata. However, these methods quickly became obsolete as bots evolved to hijack legitimate profile semantics. Consequently, recent state-of-the-art architectures have pivoted to Graph Neural Networks (GNNs). By framing bot detection as a node classification problem over a heterogeneous interaction graph, modern frameworks successfully integrate local semantic features with multi-relational topologies (e.g., ``follows'', ``retweets'', ``mentions'').

Despite these architectural advancements, deployed GNN detection systems suffer from two critical, structural vulnerabilities. The first is \textbf{Heterophily Blindness}. Standard message-passing GNNs operate under the strict assumption of \textit{homophily}---the sociological principle that similar nodes connect to one another. While early-generation bots formed dense, homophilic ``bot-nets'', modern sophisticated bots engage in \textit{relation camouflage}. By deliberately following and interacting with massive numbers of genuine human users, they create a highly heterophilic local neighborhood. During the standard GNN aggregation phase, this camouflage mathematically dilutes the bot's anomalous signature with the benign features of its human neighbors, a phenomenon known as over-smoothing by camouflage.

The second vulnerability is the \textbf{Explainability Gap}. Advanced GNNs function as impenetrable black boxes, yielding binary classifications without rationale. In real-world platform moderation, administrators cannot unilaterally suspend thousands of accounts based solely on an opaque probability score; doing so risks severe public backlash and regulatory scrutiny. Moderators require concrete, interpretable forensic evidence---such as specific suspicious interaction subgraphs or coordinated community motifs---detailing exactly \textit{why} an account was flagged. 

To simultaneously address the dual challenges of adversarial camouflage and black-box opacity, we propose \textbf{XHBot} (eXplainable Heterophily-aware Bot detector). XHBot fundamentally reimagines the aggregation process by filtering adversarial edges before message passing and maintaining strict channel separation between homophilic communities and heterophilic camouflage. Furthermore, it tightly couples detection accuracy with a hierarchical framework for transparent evidence generation.

In summary, the primary contributions of this paper are fourfold:
\begin{itemize}
    \item \textbf{Heterophily-Aware Architecture:} We introduce a novel Spectral-Guided Topology Refinement (SGTR) module and a Tri-Channel Heterophily-Aware Aggregation (THCA) mechanism that effectively neutralizes relation camouflage and prevents the mathematical dilution of bot signals.
    \item \textbf{Latent Disentanglement:} We design a Contrastive Prototype Disentanglement (CPD) objective that forces the model to decouple an account's intrinsic behavioral signature from the adversarial noise of its social positioning.
    \item \textbf{Hierarchical Forensic Explainability:} We propose an integrated Hierarchical Forensic Explanation (HFE) module that extracts interpretable evidence at the instance (subgraph) and macro (community motif) levels, empowering transparent platform moderation.
    \item \textbf{State-of-the-Art Performance:} Extensive empirical evaluations on highly camouflaged interaction graphs demonstrate that XHBot achieves an F1 score of 0.9474, outperforming state-of-the-art GNN baselines by over 9.6\% while providing full interpretive transparency.
\end{itemize}