\section{Related Works}
\label{sec:related}

\subsection{Social Bot Detection: Evolution and Recent Advances}

Social bots are automated accounts that mimic human behaviour on online platforms
to spread misinformation, manipulate public opinion, and conduct coordinated
fraud~\cite{cresci2020decade}.
A decade of research has documented an arms race between increasingly sophisticated
bots and progressively more capable detectors~\cite{cresci2020decade}.
Early bots were detectable by simple metadata rules, but third-generation
\emph{social spambots}~\cite{cresci2017paradigm} camouflage themselves at the
individual level while acting in coordinated groups, rendering per-user classifiers
ineffective and motivating the shift to graph-based collective analysis.

The introduction of large-scale graph benchmarks transformed the field.
TwiBot-20~\cite{feng2021twibot20} provided the first benchmark with explicit
follow-graph structure among 229{,}573 users, enabling relational GNN methods.
TwiBot-22~\cite{feng2022twibot22} extended this to approximately one million users
with a full heterogeneous schema covering four node types (user, tweet, hashtag,
list) and six edge types (follow, post, mention, retweet, like, contain),
re-implementing 35 baselines and revealing that no single method generalises
across all dataset domains~\cite{feng2022twibot22}.

Recent 2025--2026 work confirms that \emph{heterophily}---the tendency of bots to
connect extensively to human accounts as a camouflage strategy---is the dominant
structural challenge in bot graphs.
NeighborSense~\cite{li2026neighborsense} introduces a dynamically maintained
shortcut module in a relational GCN that adapts aggregation to local feature
distributions and heterogeneous relational structures, achieving consistent accuracy
gains over BotRGCN and RGT on both TwiBot-20 and TwiBot-22.
CB-MTE~\cite{cheng2025cbmte} fuses DistilBERT text embeddings, graph embeddings,
and manifold-learned structural features via a CatBoost-based collaborative
reasoning mechanism and validates cross-topic generalisation across five social
domains on TwiBot-22.
HW-GNN~\cite{liu2025hwgnn} takes a spectral approach, observing that homophilic and
heterophilic communities in bot networks exhibit distinct frequency distributions,
and introduces learnable Gaussian-window polynomial basis functions that target
bot-specific spectral bands, achieving an average improvement of 4.3\% in F1
across multiple benchmarks.
At the community level, \cite{fu2026heterogeneous} combine multi-dimensional
interaction graphs with explicit class-imbalance mitigation via sample balancing,
addressing both the heterogeneous relational structure and the severe bot-prevalence
imbalance jointly.
\cite{he2025bothp} propose BotHP, a dual-encoder framework evaluated at
KDD~2025 that separates ego and neighbour embeddings to mitigate interaction
camouflage while leveraging self-supervised pre-training for generalisation.

Despite these advances, \emph{no} existing work applies post-hoc subgraph
explanation to a GNN-based social bot detector evaluated on TwiBot-20 or TwiBot-22,
leaving a critical gap for forensic and platform-moderation applications.

\subsection{GNN-Based Fraud Detection}

The broader literature on GNN-based graph fraud detection provides a rich
methodological toolkit that directly transfers to bot detection.
CARE-GNN~\cite{dou2020caregnn} is the foundational camouflage-resistant fraud
detector: it introduces a label-aware similarity measure to identify informative
neighbours, a reinforcement-learning-based neighbour selector, and relation-aware
aggregation to resist feature and relation camouflage on the YelpChi and Amazon
review fraud benchmarks.
Its core insight---that fraudsters strategically connect to benign nodes to dilute
the aggregation signal---maps precisely to the bot-graph heterophily problem.

The dual-channel paradigm was recently formalised for fraud detection by
DHMP~\cite{zhang2025dhmp}, which divides the neighbourhood into homophilic and
heterophilic subgraphs and applies frequency-aware message passing independently to
each, demonstrating that separating high- and low-frequency signals substantially
improves detection on e-commerce and social-network fraud datasets.
On the efficiency and scalability front, DRHGNN~\cite{li2025drhgnn} introduces
relation-aware computation graph pre-processing and stochastic projection to reduce
redundant message passing in heterogeneous fraud graphs, making heterogeneous GNNs
practical at large scale.

The relational GCN (R-GCN)~\cite{schlichtkrull2018rgcn} remains the canonical
backbone for multi-relational graphs: it applies separate weight matrices per
relation type before summing contributions, and serves as the foundation for
BotRGCN~\cite{feng2021botrgcn} in the bot detection context and for many fraud
detection models.

An empirical study directly relevant to our design choices is presented by
\cite{dang2022evaluating}, who evaluate the contribution of each relation type
to GNN-based fraud detection on the Amazon-Fraud and Yelp-Fraud benchmarks.
Using a permutation-importance workflow---repeatedly replacing one relation's
adjacency with uniform random values and re-evaluating AUC and F1---they find
that removing any single relation type, or even all three simultaneously, causes
only a marginal drop in predictive performance, while individual node-based
features account for the dominant share of model accuracy.
This result has a direct implication for XHBot: it motivates our
Spectral-Guided Topology Refinement module (SGTR), which scores and
soft-prunes low-information camouflage edges \emph{before} aggregation rather
than treating all relations uniformly, ensuring that the tri-channel aggregation
in Module 2 operates on a cleaner graph where structurally informative edges
are weighted more heavily than noisy or camouflage-driven connections.

\subsection{GNN-Based Bot Detection with Graph Neural Networks}

BotRGCN~\cite{feng2021botrgcn} was the first to apply R-GCN to Twitter bot
detection, constructing a heterogeneous follow graph and fusing semantic, property,
and neighbourhood modalities.
RGT~\cite{feng2022rgt} replaced R-GCN's aggregation with multi-relation transformer
attention and a semantic attention network across relation types, demonstrating that
modelling relation heterogeneity yields consistent improvements on TwiBot-20.
While both models advance the state of the art, they assume homophily in their
aggregation process and provide no mechanism for handling camouflage-by-dilution,
the dominant failure mode identified in recent work~\cite{li2026neighborsense,
liu2025hwgnn}.

The work of ~\cite{qiao2026mgdil} proposes MGDIL, a unified framework for cross-domain social
bot detection that addresses two persistent limitations of prior work: brittleness
to incomplete or missing user modalities, and poor generalisation to out-of-distribution
bot populations.
MGDIL first converts heterogeneous user signals---profile metadata, historical posts,
and affective cues---into a unified textual representation via LLM-based
multi-granularity summarisation across five dimensions (content themes, sentiment
polarity, emotional tone, linguistic style, and communicative function).
Building on this representation, it applies LoRA-based instruction tuning of a
Qwen2.5-1.5B backbone, domain-adversarial training with a gradient reversal layer
to suppress dataset-specific signals, and cross-domain contrastive learning to
promote intra-class compactness and inter-class separation across domains.
A relation-aware GNN module initialised from the LLM embeddings further enriches
representations with structural context.
Trained on 13 source datasets and evaluated on the two most recent held-out datasets
(TwiBot-2020 and Fox-2023), MGDIL substantially outperforms GNN baselines including
BotRGCN and RGT under distribution shift.
While MGDIL advances cross-domain robustness, it provides no explanation for
individual bot classifications and does not model heterophily within the GNN
aggregation itself---gaps that our proposed XHBot framework directly targets.

\subsection{Explainability in Graph Neural Networks}

GNNExplainer~\cite{ying2019gnnexplainer} is the foundational instance-level
explainer: it formulates explanation as maximising mutual information between a GNN's
prediction and a learned edge-mask subgraph, and provides both structural and
feature-level explanations.
PGExplainer~\cite{luo2020pgexplainer} addresses GNNExplainer's limitation of
per-instance optimisation by parameterising the explanation generation process with
a shared deep neural network, enabling inductive, multi-instance explanations with
up to 24.7\% improvement in AUC on graph classification tasks.

Despite this active explainability literature, neither GNNExplainer nor PGExplainer
has been applied to social bot detection on TwiBot-scale graphs.
Existing bot detection models report attention weights as proxy explanations, but
these are insufficient for platform operations: moderators require \emph{which
subgraph of accounts and which relational patterns} drove a bot classification,
not scalar attention coefficients.
This explainability deficit motivates the forensic explanation module of our proposed
XHBot framework, which integrates post-hoc subgraph explanation directly with the
bot-signature disentangled representation rather than the raw GNN embedding.

RABot \cite{zhang2026rabot} is a multi-granularity graph-augmentation detector built around two practical failure modes that most GNN-based approaches leave unresolved: class imbalance arising from the scarcity of labelled bot accounts, and topological noise introduced when bots deliberately force follow-links to benign users. RABot first extracts four user-level feature channels--numerical profile attributes, Boolean metadata, free-text profile descriptions, and tweet sequences--encoding the latter two with a pretrained language model and fusing all four modalities through multi-head self-attention into a unified per-node embedding. To counter class imbalance, a neighborhood-aware oversampling module generates new bot nodes by interpolating between existing bot embeddings in the latent space, a choice that avoids the artificial edge construction required by feature-space approaches. To suppress topological noise, a reinforcement-guided edge-filtering module assesses the reliability of each edge at every GNN layer using a learned similarity score; edges that fall below a trainable threshold $\tau$ are discarded, while $\tau$ itself is continuously adjusted by a lightweight RL agent that monitors training dynamics, removing the need for manual per-dataset tuning. The denoised graph and balanced node set are then fed into the backbone GNN. RABot is architecture-agnostic, demonstrating consistent improvements over GCN, GAT, RGT, and BotRGCN backbones across Cresci-15, TwiBot-20, and MGTAB. Despite these contributions, RABot neither generates interpretable evidence for individual classifications nor explicitly accounts for the heterophilic aggregation distortion that camouflaged bots induce--two aspects that XHBot directly targets through its Tri-Channel Heterophily-Aware Aggregation mechanism and Hierarchical Forensic Explanation module.

FEDRIO \cite{yang2026fedrio} addresses a largely overlooked axis of the bot detection problem: the absence of effective privacy-preserving collaboration across platforms. Rather than improving single-platform detection, FEDRIO frames bot detection as a heterogeneous federated learning problem in which multiple social media platforms jointly train a shared detector without exchanging raw user data. At the model level, each client is equipped with an adaptive message-passing backbone comprising two cooperative networks — an action network implemented with two separate GraphSAGE modules that independently learn per-node probability distributions over message reception and emission modes, and an environment network (GIN) that performs the actual aggregation over the resulting filtered adjacency matrix, enabling node-level personalized representations across structurally heterogeneous platforms. Cross-platform knowledge transfer is achieved through a federated adversarial contrastive knowledge distillation mechanism: a shared global generator captures distributional knowledge and conveys it to each client without requiring raw data sharing, while a multi-stage adversarial training strategy with two classifiers and cross-client contrastive learning jointly align feature spaces across clients to prevent model drift. Client-side parameter updates are governed by a reinforcement learning agent formulated as a Markov Decision Process, which determines the weighted contribution of global model parameters to each local model update. Server-side aggregation is refined through a trainable neuron-level parameter mask matrix that performs element-wise weighting of each client's parameter updates during global model fusion. Evaluated on non-IID partitions of both TwiBot-20 and Vendor-19, FEDRIO consistently outperforms state-of-the-art heterogeneous federated learning baselines in detection accuracy and convergence speed. Despite its novel federated perspective, FEDRIO operates under the assumption of homophilic message passing within each client's local graph, offering no mechanism to handle heterophilic aggregation distortion caused by camouflaged bots deliberately connecting to benign users — nor does it provide instance-level forensic explanations for individual classifications — both of which are central contributions of the proposed XHBot framework.