\section{Conclusion}

In this paper, we introduced XHBot, an eXplainable Heterophily-aware Graph Neural Network designed specifically to combat advanced relation camouflage in social bot networks. By acknowledging that modern bots intentionally violate homophily to evade detection, we engineered a framework that explicitly decouples behavioral signatures from adversarial social positioning. 

Through Spectral-Guided Topology Refinement (SGTR) and Tri-Channel Heterophily-Aware Aggregation (THCA), XHBot successfully prevents the mathematical dilution of bot signals. Furthermore, the Contrastive Prototype Disentanglement (CPD) module ensures robust representation learning, while the Hierarchical Forensic Explanation (HFE) module provides the multi-level transparency—measured through high Fidelity and Sparsity—required by platform administrators.

Extensive evaluations across three distinct benchmarks—TwiBot-20, TwiBot-22, and Cresci-2017—demonstrate that XHBot achieves a new State-of-the-Art performance, consistently outperforming recent advanced baselines like HW-GNN and NeighborSense. Future work will explore adapting the XHBot disentanglement mechanisms to continuous-time dynamic graphs, enabling the real-time detection and explanation of rapidly evolving botnet camouflage strategies.
\section{Conclusion}
In this paper, we presented XHBot, a novel graph neural network framework specifically tailored to counter social bot relation camouflage. By integrating Spectral-Guided Topology Refinement and Tri-Channel Heterophily-Aware Aggregation, our model successfully decouples human noise from bot anomaly signatures. Crucially, the addition of the Hierarchical Forensic Explanation module ensures that detections are paired with actionable, interpretable evidence. Empirical evaluations demonstrate that XHBot significantly outperforms traditional baseline models, securing a massive +27.37\% F1 score improvement in challenging high-heterophily scenarios.
